---
title: "Spring 2024 Election Results"
author: "Slater Dixon"
format: 
  pdf:
    fig-pos: 'h'
editor: visual
execute:
  echo: false
  warning: false
bibliography: references.bib
---

# Voting

Per the ASA Constitution and Election Code, 2024 Spring Elections were conducted using Ranked Choice Vote, also known as Single Transferable Vote. Results were tabulated using the R package `vote`. [-@sevcikova2023]

```{r setup}

source("tabulate_results.R")
source("plot_results.R")


```

## President/Vice President

Annie Johnston was elected President and Prince Adhikari was elected Vice President by default.

## Secretary

Carson Steichen was elected Secretary by default.

## At-Large Senator

### Results

```{r}

at.large.results <- rcv_results()


get_results_table(at.large.results)
```

### Tabulation

```{r}

get_results_plot(at.large.results) |> print()

```

```{r}

at.large.results |>
  rank_distribution_plot(position = "large_senator")
```

## Second-Year Senator

### Results

```{r}
second.year.results <- rcv_results(position = "second_senator")

get_results_table(second.year.results)
```

### Tabulation

```{r}

get_results_plot(second.year.results, position = "second_senator") |> print()
```

```{r}

at.large.results |>
  rank_distribution_plot(position = "second_senator")
```

## Third-Year Senator

### Results

```{r}

third.year.results <- rcv_results(position = "third_senator")

get_results_table(third.year.results)
```

### Tabulation

```{r}

get_results_plot(third.year.results, position = "third_senator") |> print()
```

```{r}

at.large.results |>
  rank_distribution_plot(position = "third_senator")
```

## International Senator

### Results

```{r}

international.results <- rcv_results(position = "international_senator", seats = 2)

get_results_table(international.results)
```

### Tabulation

```{r}

get_results_plot(international.results, position = "international_senator")
```

```{r}

at.large.results |>
  rank_distribution_plot(position = "international_senator")
```

## Fourth-Year Senator

The following candidates were elected by default:

-   Isha Hooda

-   Raleigh Lunderman

-   Avery Gilman

-   Domanic Dick

    ```{r}

    fourth.year.results <- rcv_results(position = "fourth_senator", seats = 3)

    candidates <- colnames(fourth.year.results$data)
    winners <- fourth.year.results$elected


    ordered <- c(
      candidates, 
      winners[which(!candidates %in% winners)]
    )


    fourth.year.results |>
      rank_distribution_plot(position = "fourth_senator", bold.winners = F)

    ```

# Turnout

```{r}

turnout.df <- read_csv("./data/sp24_turnout_results.csv", col_select = c("Race", "Votes", "Eligible")) |> 
  head(7) |>
  mutate(across(c(Votes, Eligible), ~as.integer(.))) 


turnout.df |>
  filter(!(grepl("President/VP*|Secretary*", Race))) |>
  mutate(Race = str_replace(Race, "At-Large", "At-Large/Pres./Sec.*")) |>
  mutate(Turnout = paste0(((Votes / Eligible) * 100)|> round(2), "%")) |>
  kbl() |>
    kable_material("striped") |>
  footnote(symbol = "Uncompetitive")
```

## By Race

```{r}

turnout.df |>
  filter(!(grepl("President/VP*|Secretary*", Race))) |>
  rename(Voted = Votes) |>
  mutate(Race = str_replace(Race, "At-Large", "At-Large/Pres./Sec.")) |>
  mutate(`Did Not Vote` = Eligible - Voted) |>
  select(-Eligible) |>
  pivot_longer(-Race) |>
  arrange(desc(value)) |>
  ggplot() +
  geom_bar(aes(x = Race, y = value, fill = name),  stat="identity", position="fill") +
  scale_fill_manual(values = c("Did Not Vote"  = "#273046", "Voted" = "#E54E21")) +
    coord_flip() +
  theme_minimal() +
  labs(
    title = "ASA SP24 Elections Voter Turnout by Race",
    fill = "",
    x = "Race",
    y = "Percentage"
  ) +
    theme_tufte() +
    theme(plot.title = element_text(hjust = 0.5, face = "bold"),
          plot.subtitle = element_text(hjust = 0.5),
          axis.text.y = ggtext::element_markdown(lineheight = .25),
          plot.caption = ggtext::element_markdown(),
          legend.position = "bottom",
          strip.text.x = element_text(
            size = 12, color = "#404040", face = "italic"
          ))
  
```

## Historical

```{r}

library(janitor)

read_csv("./data/asa_historical_turnout.csv") |>
  head(8) |>
  mutate(`Competitive President?` = if_else(as.logical(`Competitive President?`), "", "*"),
         `Election Year` = paste0(`Election Year`, `Competitive President?`)) |>
  select(c("Election Year", "Votes Cast", "Total Fall Undergraduate Enrollment")) |>
  rename(Voted = `Votes Cast`) |>
  mutate(`Did Not Vote` =  `Total Fall Undergraduate Enrollment` - Voted) |>
  select(-`Total Fall Undergraduate Enrollment`) |>
  pivot_longer(-`Election Year`) |>
  arrange(desc(`Election Year`)) |>
  ggplot() +
  geom_bar(aes(x = `Election Year`, y = `value`, fill = name), stat = "identity", position = "fill")  +
  scale_fill_manual(values = c("Did Not Vote"  = "#273046", "No Data" = "#404040", "Voted" = "#E54E21")) +
   # coord_flip() +
  labs(
    title = "ASA Spring Elections Historical Turnout",
    x = "Race",
    y = "Percentage",
    caption = "*Uncompetitive President/VP race",
    fill = ""
  ) +
theme_tufte() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
          plot.subtitle = element_text(hjust = 0.5),
          axis.text.y = ggtext::element_markdown(lineheight = .25),
          plot.caption = ggtext::element_markdown(),
          legend.position = "bottom",
          strip.text.x = element_text(
            size = 12, color = "#404040", face = "italic"
          ))
```
